Approximation as Differentiation

### 2025-05-07 main/formal-structures.tex
Addition of morphisms derived from biproduct structure.

>>>>

\begin{definition}[Biproducts and Semi-additive categories]
A category with finite products and coproducts is \emph{semi-additive} if the canonical morphisms (projections
and inclusions) give an isomorphism
\[\textstyle X \coprod Y \iso X \prod Y\] that is natural in both variables. The product/coproduct is called a
\emph{biproduct}, and the biproduct structure is denoted by $(\biprod, 0)$. The unit $0$ is both terminal and
initial and is called a \emph{zero} object.
\end{definition}

A semi-additive category $\cat{C}$ is enriched in $\CMon$, the category of commutative monoids: for any two
morphisms $f, g: X \to Y$ in $\cat{C}$, the biproduct structure provides a way to ``add'' them together,
forming a morphism $f + g: X \to Y$. Diagrammatically:

\begin{center}
\begin{tikzcd}
   X \arrow[r, "\diag"] & X \biprod X \arrow[r, "f \biprod g"] & Y \biprod Y \arrow[r, "\codiag"] & Y
\end{tikzcd}
\end{center}

\noindent Here $\diag$ denotes the diagonal $\prodM{\id_X}{\id_X}$ given by the universal property of the
product and $\codiag$ denotes the codiagonal $\coprodM{\id_X}{\id_Y}$ given by the universal property of the
coproduct. The $f \oplus g$ morphism is the component-wise map $\prodM{f}{g} = \coprodM{f}{g}$.
